---
title: "Hands-on Exercise 1: Geospatial Data Wrangling with R"
execute: 
  warning: false
  freeze: true
date: "`r Sys.Date()`"
---

# Overview

Welcome to my first hands-on exercise of this module, IS415 Geospatial and Applications!

In this exercise, our primary aim is to delve into geospatial data science tasks using the capabilities of the sf package in R.

-   Installing and loading sf and tidyverse packages into R environment

-   Importing geospatial and aspatial data by using appropriate functions (sf and readr package)

-   Exploring the content of simple feature dataframe

-   Assigning and transforming coordinate systems

-   Performing geoprocessing, data wrangling and Exploratory Data Analysis (EDA) tasks

# Getting Started

## Installing and loading R packages

In this section, I will install and load tidyverse and sf packages.

```{r}
pacman::p_load(tidyverse, sf)
```

# Importing geospatial data

## Importing polygon feature data in shapefile format

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

## Importing polyline feature data in shapefile form

```{r}
cyclingpath <- st_read(dsn = "data/geospatial", 
                  layer = "CyclingPathGazette")
```

## Importing GIS data in kml format

```{r}
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```

# Checking the content of a simple feature data frame

```{r}
st_geometry(mpsz)
glimpse(mpsz)
head(mpsz, n=5)  
```

# Plotting the Geospatial Data

```{r}
#| echo: false
plot(mpsz)
```

## Plotting the geometry

```{r}
#| echo: false
plot(st_geometry(mpsz))
```

## Plotting a specific attribute eg. "PLN_AREA_N"

```{r}
#| echo: false
plot(mpsz["PLN_AREA_N"])
```

# Assigning EPSG code to a simple feature data frame

```{r}
st_crs(mpsz)
```

## Assigning the correct EPSG code (3414) to mpsz dataframe

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
```

## Transforming the projection of preschool from WGS84 to SVY21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
st_crs(preschool3414)
```

# Importing the aspatial data

```{r}
#| echo: false
listings <- read_csv("data/aspatial/listings.csv")
list(listings) 
```

## Creating a simple feature dataframe from an aspatial dataframe

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

## Examining the contents of the newly created simple data frame

```{r}
#| echo: false
glimpse(listings_sf)
```

# Geoprocessing with sf package

## Buffering

Buffer of 5 metres of reserved land on the both sides of the current cycling path

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

## Point-in-Polygon Count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
```

## Density of Preschools by planning subzone

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# Exploratory Data Analysis (EDA)

```{r}
#| echo: false
hist(mpsz3414$`PreSch Density`)
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
