---
title: "Hands-on Exercise 3"
format: 
  html:
    code-fold: true
    code-summary: "Show the code"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
date: "`r Sys.Date()`"
---

# Data

| Data Set                  | Description                                                                                       | Source                                           |
|----------------|--------------------------------------|-------------------|
| [`CHILDCARE`]{.smallcaps} | Point feature data providing both location and attribute information of childcare centres         | data.gov.sg (geojson format)                     |
| `MP14_SUBZONE_WEB_PL`     | Polygon feature data providing information of URA 2024 Master Plan Planning Subzone boundary data | data.gov.sg (ESRI shapefile format)              |
| `CoastalOutline`          | Polygon feature data showing the national boundary of Singapore                                   | Singapore Land Authority (ESRI shapefile format) |

# Installing and Loading the R packages

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

# Spatial Data Wrangling

## Importing the spatial data

```{r}
childcare_sf <- st_read("data/geospatial/ChildCareServices.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "data/geospatial", layer="CostalOutline")
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

## Transforming the data

```{r}
st_geometry(mpsz)
mpsz_sf <- st_set_crs(mpsz,3414)
mpsz_sf
```

```{r}
sg_sf
```

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

# Geospatial Data Wrangling

## Converting sf dataframes to sp's spatial class

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

Inspecting the information

```{r}
childcare
mpsz
sg
```

## Converting the spatial class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
childcare_sp
sg_sp
```

## Converting the genric sp format into spatstat's ppp format

```{r}
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
#plot(childcare_ppp)
```
